\name: A Non-Parametric, Hierarchical \\ Wavelength-Calibration Method for a Precision Spectrograph

### 2020-07-16 paper/paper.tex
Pg1 Hogg Edits

\begin{document}
\title{\name:
  A Non-Parametric, Hierarchical \\
  Wavelength-Calibration Model for a Precision Spectrograph}

@@

\begin{abstract}\noindent%
\name\ is a hierarchical, non-parametric framework for precision wavelength-calibration of spectrographs.  This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.  Instruments are likely to vary along only a few dominant degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.  \name\ takes advantage of this property by using all calibration data to construct a low-dimensional space of all possible calibrations for an instrument.  \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.  This density of calibration lines permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.  We demonstrate the success of this method with data from the EXtreme PREcision Spectrograph, \expres, which includes a Menlo Systems laser frequency comb.  Using \name\ wavelengths to predict the wavelengths of known calibration lines returns smaller per-line RMS by a factor of 5 when compared to a classic wavelength model.  Employing \name\ wavelengths on \expres\ RV measurements of HD 3651 reduced the RMS of a planet fit by \lz{blah}.
\end{abstract}

@@

\section{Introduction} 
Precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.  These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.  Exoplanet science happens at the level of $3\,\mps$ precision, which is sensitive enough to discover gas-giant planets around other stars.  The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds \citep{fischer2016}.  This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).    Some hardware--software systems, such as \espresso, \expres,  \neid, etc., were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.  Early data from \expres\ and \espresso\ suggest that the hardware improvements truly propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, mascareno2020}.

@@

\subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
We tested the effect of \name -generated wavelengths on RV measurements using \expres\ observations of \target.  \target\ is a BLANK star and also SOME STATEMENT ON SUSPECTED ACTIVITY(?).  We used BLANK observations of \target\ taken between BLANK and BLANK with an average signal-to-noise ratio of BLANK PROBABLY 250.  Radial velocity mesurements are derived using the forward-modeling scheme employed by \expres's pipeline \citep{petersburg2020}.

@@

\section{Choose Your Own Implementation} \label{sec:choices}
We have described and tested only one implementation of \name.  Using PCA and an interpolated wavelength solution is a statisticall straight-forward step towards a complete hierarchical and non-parametric wavelength model.  It is possible to upgrade both the denoising and wavelength solution steps to true models.

>>>>

\begin{document}
\title{\name:
  A Non-Parametric, Hierarchical \\
  Wavelength-Calibration Method for a Precision Spectrograph}

@@

\begin{abstract}\noindent%
\name\ is a non-parametric, hierarchical framework for precision wavelength-calibration of spectrographs.  This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.  Instruments are likely to vary along only a few dominant degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.  \name\ takes advantage of this property by using all calibration data to construct a low-dimensional space of all possible calibrations for an instrument.  \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.  This density of calibration lines permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.  We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph}, \expres, which includes a laser frequency comb.  Using \name\ wavelengths to predict the wavelengths of known calibration lines returns smaller per-line RMS by a factor of 5 when compared to a classic wavelength solution that is constructed using polynomials and treats each exposure individually .  Employing \name\ wavelengths on \expres\ RV measurements of HD 3651 reduced the RMS of a planet fit by \lz{blah}.
\end{abstract}

@@

\section{Introduction} 
Precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep[e.g.][]{mayor2011, bonfils2013, plavchan2015, butler2017}.  These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.  Exoplanet science happens at the level of $3\,\mps$ precision, which is sensitive enough to discover gas-giant planets around other stars.  The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds \citep{fischer2016}.  This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).    Some hardware--software systems, such as \espresso, \expres,  \neid, etc., were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.  Early data from \expres\ and \espresso\ suggest that the hardware improvements truly propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, mascareno2020}.

@@

\subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
We tested the effect of \name -generated wavelengths on RV measurements using \expres\ observations of \target.  \target\ is a BLANK star and also SOME STATEMENT ON SUSPECTED ACTIVITY(?).  We used BLANK observations of \target\ taken between BLANK and BLANK with an average signal-to-noise ratio of BLANK PROBABLY 250.  Radial velocity mesurements are derived using a chunk-by-chunk, forward-modeling algorithm employed by \expres's pipeline \citep{petersburg2020}.

@@

\section{Choose Your Own Implementation} \label{sec:choices}
We have described and tested only one implementation of \name.  Using PCA and an interpolated wavelength solution is a statisticall straight-forward step towards a complete hierarchical and non-parametric wavelength model.  It is possible to upgrade both the denoising and wavelength solution steps to true models.  It is also possible, of course, to implement either step individually.  A hierarchical framework can be used to simply denoise the lines before they are fit to a parametric model, or use a non-parametric model on lines that haven't been denoised.
